---
title: "Introduction to Programming Concepts"
author: "Nikki Rubinstein"
date: "Saturday, June 25, 2016"
output: ioslides_presentation
runtime: shiny
--- 

```{r, include=FALSE}
library(shiny)
library(jpeg, lib.loc = "lib")
library(png, lib.loc = "lib")
library(grid)
```

## Telling your computer what to do
<!-- concept of programming languages -->
<!-- play simon says -->
```{r fig.width=5, fig.height=5,echo=FALSE, fig.align='center'}
img <- readJPEG("fig//Olate_dogs_-_on_wheel.jpeg")
grid.raster(img) 
``` 

<!-- image source: https://upload.wikimedia.org/wikipedia/commons/d/d8/Olate_dogs_-_on_wheel.jpeg -->

## Telling your computer what to do

Some programming languages you might have heard of:  

* R
* Matlab
* Python
* JavaScript
* html

## Interacting with your computer
<!-- introduce the idea of variables, assignment and memory storage
at this point people should name their teams / tables -->
```{r fig.width=5, fig.height=5,echo=FALSE, fig.align='center'}
img <- readPNG("fig//Abstract_art_(PSF).png")
grid.raster(img) 
``` 
<!-- image source: https://upload.wikimedia.org/wikipedia/commons/6/66/Abstract_art_(PSF).png -->

## Interacting with your computer

```{r fig.width=5, fig.height=5,echo=FALSE, fig.align='center'}
img <- readJPEG("fig//pexels-photo.jpg")
grid.raster(img) 
``` 
<!-- image source: https://static.pexels.com/photos/15393/pexels-photo.jpg -->

## Interacting with your computer
Creating variables

```{r}
baby <- 'Meredith'
sculpture <- 10
```
```{r}
baby 
sculpture
```

## What types of information does your computer understand?
<!-- These are the building blocks of our conversation with the computer, much like words are the building blocks of a book -->
* **numeric** 
    + integers, e.g. `1, 2, 3`
    + doubles, e.g. `1.25, 3.78`
    + complex numbers, e.g. `3 + 2i`
* **character or string** 
    + e.g. `"Research", "Platforms"`
* **logical or Boolean** 
    + e.g. `TRUE` and `FALSE`

## How can we group information?



<!-- much like how we create sentences from words -->
* vectors / tuples
* matrices / data.frames / arrays
* lists / structures
* dictionaries

## vectors / tuples
```{r fig.width=5, fig.height=5,echo=FALSE, fig.align='center'}
img <- readJPEG("fig//lechner-50119_960_720.jpg")
grid.raster(img) 
``` 
<!-- image source: https://pixabay.com/static/uploads/photo/2012/06/16/20/52/lechner-50119_960_720.jpg -->

## vectors / tuples 
```{r}
vector1 <- 1:10
vector2 <- c("learning", "to", "code", "is", "fun")
```
```{r}
vector1
vector2
```

## matrices / data.frames / arrays
```{r fig.width=5, fig.height=5,echo=FALSE, fig.align='center'}
img <- readJPEG("fig//House_of_Air_The_Matrix.jpg")
grid.raster(img) 
``` 
<!-- image source: https://upload.wikimedia.org/wikipedia/commons/f/fa/House_of_Air_The_Matrix.jpg -->

## matrices / data.frames / arrays
```{r}
PlatoonLeads <- data.frame(
    platoon = c("Data Wranglers", 
                "Data Miners", 
                "Data Vizards", 
                "Cadventurers"),
    name    = c("Kerry Halupka", 
                "Kim Doyle", 
                "Isabell Kiral-Kornek", 
                "Louise van der Werff"))
```
```{r}
PlatoonLeads
```
## lists / structures
```{r fig.width=5, fig.height=5,echo=FALSE, fig.align='center'}
img <- readJPEG("fig//musical instrument collection.jpg")
grid.raster(img) 
``` 
<!-- image source: http://s0.geograph.org.uk/geophotos/03/54/79/3547925_df1f2e83.jpg -->

## lists / structures
```{r}
shoppingList <- list(
  breakfast = c("cereal", "milk", "orange juice", "banana"),
  lunch     = c("bread", "cheese", "tomato"),
  dinner    = c("frozen pizza", "chocolate mousse")
  )
```
```{r}
shoppingList
```
## dictionaries
```{r fig.width=5, fig.height=5,echo=FALSE, fig.align='center'}
img <- readJPEG("fig//dictionary.jpg")
grid.raster(img) 
``` 
<!-- image source: https://pixabay.com/static/uploads/photo/2015/12/23/22/16/collaboration-1106196_960_720.jpg -->

## dictionaries
```{python}
IMDBRatings = {"Game of Thrones": 9.4, 
    "Sherlock": 9.2, 
    "Firefly": 9.1, 
    "Friends": 8.9}

for movie, rating in IMDBRatings.items():
    print movie + ': ' + str(rating)
```
## How to add flexibility to the conversation
**If statements**
```{r fig.width=5, fig.height=5,echo=FALSE, fig.align='center'}
img <- readJPEG("fig//Am I hungry.jpg")
grid.raster(img) 
``` 

## How to add flexibility to the conversation
**If statements**
```{r,echo=F}
selectInput("hungry", "Am I hungry?", choices = c(TRUE, FALSE))
```
```{r,eval=F}
if (hungry){
  print("GIVE ME A TIM TAM!!!")
} else {
  print("No tim tam for me, thanks!")
}
```
```{r,echo=F}
renderText({if (input$hungry){
  print("GIVE ME A TIM TAM!!!")
} else {
  print("No tim tam for me, thanks!")
}})
```

## How to add flexibility to the conversation
**For loops**
```{r fig.width=3, fig.height=4,echo=FALSE, fig.align='center'}
img <- readJPEG("fig//Man_and_woman_dancing_a_waltz.jpg")
grid.raster(img) 
``` 
<!-- image source: https://upload.wikimedia.org/wikipedia/commons/4/49/Man_and_woman_dancing_a_waltz_(rbm-QP301M8-1887-197a~6).jpg -->

## How to add flexibility to the conversation
**For loops**
```{r}
for (beats in 1:4){
  print("bring left foot to right foot")
  print("step left foot back to original position")
  print("bring right foot to left foot")
  print("step left foot back to original position")
}
```
## How to add flexibility to the conversation
**Functions**
```{r fig.width=5, fig.height=5,echo=FALSE, fig.align='center'}
img <- readJPEG("fig//cooking.jpg")
grid.raster(img) 
``` 
<!-- image source: https://pixabay.com/static/uploads/photo/2015/10/30/05/48/cooking-1013455_960_720.jpg -->

## How to add flexibility to the conversation
**Functions**
```{r, echo=F}
selectInput("day", "Day of the week", choices = c("Monday", "Tuesday", 
                                                  "Wednesday", "Thursday",
                                                  "Friday", "Saturday",
                                                  "Sunday"))
```
```{r}
moodCalendar <- function(day){
  if (day %in% c("Friday", "Saturday", "Sunday")){
    print(paste(day, "is a happy day"))
  } else if (day %in% c("Monday", "Tuesday")){
    print(paste(day, "is a grumpy day"))
  } else {
    print(paste(day, "is a neutral day"))
  }}
```
```{r, echo=F}
renderText({moodCalendar(input$day)})
```

## Where to find help
```{r fig.width=5, fig.height=2,echo=FALSE, fig.align='center'}
img <- readPNG("fig//stackoverflow.png")
grid.raster(img) 
```
```{r fig.width=5, fig.height=2, echo=FALSE, fig.align='center'}
img <- readJPEG("fig//google.jpg")
grid.raster(img) 
``` 

## Making mistakes is the best way to learn
```{r fig.width=7.5, fig.height=4.5,echo=FALSE, fig.align='center'}
img <- readJPEG("fig//it's ok to make mistakes.jpg")
grid.raster(img) 
``` 

## Where to learn more

* Sign up for one of our training courses: [http://melbourne.resbaz.edu.au/participate](http://melbourne.resbaz.edu.au/participate)
* [http://www.w3schools.com/](http://www.w3schools.com/)
* [MIT Open Courseware](http://ocw.mit.edu/courses/electrical-engineering-and-computer-science/6-094-introduction-to-matlab-january-iap-2010/)
* [Coursera Data Science Specialisation](https://www.coursera.org/specializations/jhu-data-science)